Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-06-03 image.tex
versions

% SWT without epsilon-transitions
% SWT do not read the other tape in "skip" transitions 10 or 01
%     (hence this version is simpler  than V2)
% SWA without epsilon-transitions
% SW-VPA read top of stack (when not empty) at every transition 
%        (6 kinds of transitions).

\documentclass[runningheads]{llncs}
%\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{hyperref}
%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
%usepackage{url}

% for footnote ref
\usepackage{refcount}

% array and tabular
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

% extension of enumerate env. (style for displaying counters)
% \usepackage{enumerate}

%% pictures
% \usepackage{graphicx} 
% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
% \graphicspath{fig/}

%% PGF, Tikz
\usepackage{tikz-cd}
%% \usepackage{pgfplots}
%% \usepgfplotslibrary{dateplot}
%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
% \usepackage{tikz}
% \usetikzlibrary{cd}
%% \usetikzlibrary{arrows}
%% \usetikzlibrary{calc}
%% \usetikzlibrary{snakes}
%% \usetikzlibrary{backgrounds}
% \usetikzlibrary{trees}
%% \usetikzlibrary{automata}
%% \usetikzlibrary{positioning}
%% \usetikzlibrary{matrix}
%% \usetikzlibrary{patterns}
%% \usetikzlibrary{shapes}

% symbols
\usepackage{amsmath} 
\allowdisplaybreaks
\usepackage{amssymb} 
\usepackage{amsbsy}
\usepackage{bbold}
\usepackage{latexsym}
%\usepackage{amsfonts}
\usepackage{stmaryrd}
%\usepackage{mathabx}
%\usepackage{MnSymbol}
\usepackage{harmony} % simple music fonts
\usepackage{mathtools} % for arrows
%\usepackage{mathptmx}

%% theorem environments
\usepackage{theorem}
% \newtheorem{theorem}{Theorem} %[section]
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{proposition}[theorem]{Proposition}
% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
% \theorembodyfont{\slshape}
% \newtheorem{example}[theorem]{Example}
% \newtheorem{remark}[theorem]{Remark}
\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

%% algorithms
%\usepackage{algorithm} 
%\usepackage{program}

%% allows for page break within arrays
\usepackage{longtable}

%% arrows etc
%\input{rewriting}

%% music symbols
% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
\usepackage{musicography}

%% for new macros
\usepackage{xspace}

%% Misc macros
\def\ie{\textit{i.e.}\xspace}
\def\eg{\textit{e.g.}\xspace}
\def\wrt{\textit{wrt}\xspace}
%\def\wlog{\textit{wlog}\xspace}
\def\etc{\textit{etc}\xspace}

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}

%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\config}[2]{\ensuremath{#1}[{#2}]} 
\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
\def\stacksep{\cdot}
\def\stackup{\\}

\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}

% \bigominus
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
<-6> mathx5 <6-7> mathx6 <7-8> matha7
<8-9> mathx8 <9-10> mathx9
<10-12> mathx10 <12-> mathx12
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}

%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}

%\sloppy

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
%\title{Symbolic Weighted Parsing and Automated Music Transcription}
%\title{Symbolic Weighted Language Models and Automated Music Transcription}
\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
\author{Florent Jacquemard}
\institute{INRIA \& CNAM, Paris, France\\
\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
%\authorrunning{Florent Jacquemard}

\date{\today}

\begin{document}
\thispagestyle{empty}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro}
\input{introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% prelin
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminary Notions}
\label{section:prelim}\label{sec:prelim}

%notations: for set $S$ : sets of sequences $S^*$ and $S^+$...
%interval $[i..j]$ of natural numbers

\subsection{Semirings} 
\label{section:semiring}\label{sec:semiring}
\input{semiring}

\subsection{Label Theory}
\label{section:symbols}
\input{labels}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SWT & SWA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Automata and Transducers}
\label{section:transducer}\label{sec:transducer}
\label{section:SWA}\label{sec:SWA}
\label{section:SWT}\label{sec:SWT}

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
between words and languages, using weighted transducers, 
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers. 
%
The models introduced in this section generalize 
weighted automata and transducers~\cite{Droste09handbook} 
%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labelling each transition with a weight function (instead of a simple weight value), 
that takes the input and output symbols as parameters. 
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an generic semiring, 
whereas the latter guards are restricted to the Boolean semiring.

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)
\noindent 
Let $\Semiring$ be a commutative semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi$ be a label theory over $\Semiring$
containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a triplet of transition functions 
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
For convenience, we shall sometimes present transition functions 
as functions of  
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}      
\]
%
\noindent 
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called input and output word.
% \noindent
More precisely, $T$ defines a mapping 
from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
based on an intermediate function $\weight_T$
defined recursively for every states $q, q' \in Q$, 
and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ by:
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
%\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
%    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
%\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
%    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}} 
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\ 
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}} 
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}} 
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}
%
We recall that, by convention (Section~\ref{sec:semiring}), 
an empty sum with $\bigoplus$ is equal to~$\zero$. 
%
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
when reading respectively $a$ and $b$ at the current positions in the input and output words, 
increment the current position in the input word if and only if $i = 1$, 
and in the output word iff $j = 1$ (otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions, 
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in 
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
symbolic transducers.

%whereas considering the current symbol may be useful to compute a transition weight.
%(even when it does not change the head's position, like with $\varepsilon$-transitions).
%
%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
%are missing in the definition of $\weight_T$.
%It means that $T$ must avoid configurations where it reached the end of 
%the output word and not of the input one, or vice-versa.
%This can be done by using $\wei_{10}$ and $\wei_{01}$
%before reaching the end of word, and using a special state for this purpose.

The expression \eqref{eq:SWT-weight} 
can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei_{10}(q, a, \varepsilon, r)$ is the cost of 
the deletion of the symbol~$a \in \Sigma$ in~$s$, 
$\wei{01}(q, \varepsilon, b, r)$ is the cost 
of the insertion of~$b \in \Delta$ in $t$, 
and $\wei_{11}(q, a, b, r)$ is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all possible products.

\medskip\noindent
%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$. 
Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is: 
%defined as follows:
\begin{equation}
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
\label{eq:SWT-value}
\end{equation}

%\begin{example}
%We can define with a \SWT the computation of a similarity measure 
%between timed sequences
%similar to dynamic time warping (DTW).
%
%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
%We consider a \SWT with one state $q$ and transitions
%$\wei_{11}(q, d, d', q) = 
% \wei_{01}(q, d, d', q) = 
% \wei_{10}(q, d, d', q) = |d' - d|$,
%for all $d, d' \in \mathbb{R}_+$.
%% needs reading input/output symbols even by epsilon-transitions.
%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
%for the computation of an optimal match between words, 
%the matching cost for two symbols being the 
%the time distance between them.
%\endex
%\end{example}

\begin{example}
In Common Western Music Notation~\cite{Gould11BB}, 
several symbols may be used to represent a unique sounding event.
For instance, several notes can be combined with a tie, 
like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$,  
and a note can be augmented with a dot like in~\musQuarterDotted).
We propose a small transducer that compares an input sequence of sounding 
events (music "performance") 
to an output sequence of written events (music "score").
% simple pointwise distance between two sequences of timestamped events **

Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
%This example of $\Delta$ is motivated by the case of music notation, 
%where several notated events (notes) can be tied together, 
%with a \emph{tie} or a \emph{dot}
%meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and 
$\init(q_1) = \final(q_1) = \zero$. 
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and ${\wei_1}$ is a transition functions 
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
%      
\noindent
As above in the case of \SWT, 
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
The computation of $A$ on words $s \in \Sigma^*$
is defined with an intermediate function $\weight_A$, 
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
%
\begin{align}
\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to 
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q') 
\label{eq:SWA-value}
\end{equation}

%
%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

%\subsection{Properties}
\noindent
The following property will be useful to the approach on 
symbolic weighted parsing presented in Section~\ref{sec:parsing}.

@@

Every non-null transition of $A_{T, s}$ will
simulate a sequence of transitions of $T$ performing the following steps:
advance in the input word while staying immobile in the output word, 
and then make one step in the output word (and advance in the input word or not).
%
The first steps in the sequence correspond to $\varepsilon$-transitions of automata, 
and their total weight is computed by the following intermediate function 
$\wei'_0 : Q' \times Q' \to \Semiring$, 
defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
by:
\[
\begin{array}{rcl}
\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
\quad \mathrm{if~} 1 \leq i \leq n+1,\\
%
\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
%
\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
\displaystyle
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
           \scriptstyle q_0 = q, q_k = q'
           \end{array}} 
\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
\quad \begin{array}[t]{l}
      \mathrm{if~} 1 \leq i < n,
      \mathrm{~and~} 1 \leq k \leq n-i,\\
      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
      \end{array}
\end{array}
\]
%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
%
The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
(Section~\ref{section:symbols}). 
% closed under partial applications 
The sum and product in its definition are finite, 
we shall see below how to compute the first sum in polynomial time.
%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
%of length $k+1$ in the sum.
%Indeed, it amounts to search short

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SW-VPA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented in this section generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.
It will compute on nested words over infinite alphabets, 
associating to every such word a weight value. 
Nest words are able to describe structures of labeled trees, 
and in the context of parsing, they will be useful to 
represent parse trees.

\subsection{Definition}\label{se:SWVPA-def}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three 
subsets~$\Omegai$, $\Omegac$, $\Omegac$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
stand respectively 
for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$ 
%of unary functions in $\Omegai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

>>>>

@@

Every non-null transition of $A_{T, s}$ will
simulate a sequence of transitions of $T$ performing the following steps:
advance in the input word while staying immobile in the output word, 
and then make one step in the output word (and advance in the input word or not).
%
The first steps in the sequence correspond to $\varepsilon$-transitions of automata, 
and their total weight is computed by the following intermediate function 
$\wei'_0 : Q' \times Q' \to \Semiring$, 
defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
by:
\[
\begin{array}{rcl}
\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
\quad \mathrm{if~} 1 \leq i \leq n+1,\\
%
\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
%
\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
\displaystyle
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
           \scriptstyle q_0 = q, q_k = q'
           \end{array}} 
\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
\quad \begin{array}[t]{l}
      \mathrm{if~} 1 \leq i < n,
      \mathrm{~and~} 1 \leq k \leq n-i,\\
      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
      \end{array}
\end{array}
\]
%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
%
The function $\wei'_0$ is defined thanks to the closure properties of the label theory~$\bar\Phi$
(Section~\ref{section:symbols}). 
% closed under partial applications 
The sum and product in its definition are finite, 
we shall see below how to compute the first sum in polynomial time.
%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
%of length $k+1$ in the sum.
%Indeed, it amounts to search short

@@
